---
author: Esa Eslami
date: 6/11/2018
title: "COMPSCIX 415.2 Homework 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL: [https://github.com/esaeslami/compscix-415-2-assignments](https://github.com/esaeslami/compscix-415-2-assignments)

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```

```{r load_data, warning=FALSE, message=FALSE}
data("WorldCities")
glimpse(WorldCities)
```

*__QUESTION 1:__ How many observations and variables are there in this data set? What are some of the variable names? Type up your answer in your document using complete sentences.*

**ANSWER 1:** There are 23,018 observations and 10 variables in the "WorldCities" data set. The 10 variable names are: code, name, latitude, longitude, country, countryRegion, population, regionCode, region, and date.

__Extract parts of the data (*EXAMPLE*)__

```{r basic extract example, warning=FALSE, message=FALSE}
WorldCities <- head(WorldCities, 200) # 200 rows
country_col <- WorldCities$country
unique(country_col)
```

*__QUESTION 2:__ There is a column called region in the data set. Can you extract this and show only the unique values?*

**ANSWER 2:** 
```{r Answer 2, warning=FALSE, message=FALSE}
region_col <- WorldCities$region
unique(region_col)
```

__Extract/select data the tidy way (*EXAMPLE*)__
```{r tidy extract example, warning=FALSE, message=FALSE}
country_col <- select(WorldCities, country)
country_col <- WorldCities %>% select(country)
WorldCities %>% select(region) %>% head(5)
```

*__QUESTION 3:__ Can you extract and show the unique entries from the country column in WorldCities using one line of code and two %>% operators? The output will look like this:*
```{r Question 3, warning=FALSE, message=FALSE}
## country
## 1 AD
## 3 AE
## 15 AF
## 65 AG
## 66 AI
## 67 AL
## 87 AM
## 104 AO
## 131 AR
```

**ANSWER 3:** 
```{r Answer 3, warning=FALSE, message=FALSE}
WorldCities %>% select(country) %>% unique()
```

__Create a visualization (*EXAMPLE*)__
```{r example histogram, warning=FALSE, message=FALSE}
WorldCities %>% ggplot(aes(x = population)) +
geom_histogram()
```

*__QUESTION 4:__ Make one more histogram of the population and add a new x-label, a new ylabel, a new title (use ggtitle('my title')), and change the theme of the plot using theme_bw().*

```{r Answer 4, warning=FALSE, message=FALSE}
WorldCities %>% ggplot(aes(x = population)) +
  geom_histogram() +
  xlab("City Population") +
  ylab("Count") +
  ggtitle('World City Populations') +
  theme_bw()
```






