---
author: Esa Eslami
date: 7/01/2018
title: "COMPSCIX 415.2 Homework 4"
output: html_document
self-contained: true
---

```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(nycflights13)
library(gridExtra)

```

* * * 

####__Homework 4 Assignments:__####
__Complete these Exercises in R for Data Science:__  

1. Section 5.6.7: #2, #4 and #6 only. Extra Credit: Do #5
2. Section 10.5: #1, #2, #3 and #6 only
3. Section 12.3.3: #2, #3 and #4 only
4. Section 12.4.3: #1 and #2 only

* * *

__Answer these questions:__

Follow these steps:

* Download the baby_names.txt file from Canvas which is in the Homework 4 assignment section.
* Load this file into R correctly and take a glimpse of the output.
* Export this file as a csv file and call it ‘baby_names.csv’.
* Reload the baby_names.csv file and take another glimpse.
* Show all of your code and the output. There should be two data import lines of code, one data export
line of code, and two glimpses of the data.

* * * 

####5.6.7 Exercises:####

*2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).*

__ANSWER:__ 

```{r 5_6_7_EX2, warning=FALSE, message=FALSE}
# First we need to create non_cancelled
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
# Now run not_cancelled %>% count(dest) to see what it looks like
not_cancelled %>% count(dest)
# We can get this same output using summarize() like this:
flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
  group_by(dest) %>%
  summarize(not_cancelled = sum(!is.na(dep_delay)))
# Confirm that the output is identical:
countdest1 <- not_cancelled %>% count(dest)
countdest2 <- flights %>%
        filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
        group_by(dest) %>%
        summarize(not_cancelled = sum(!is.na(dep_delay)))
identical(countdest1$n,countdest2$not_cancelled)
# Similarly, we can use summarize() to get the same results from not_cancelled %>% count(tailnum, wt = distance) like this:
taildist1 <- not_cancelled %>% count(tailnum, wt = distance)
taildist2 <- flights %>%
        filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
        group_by(tailnum) %>%
        summarize(not_cancelled = sum(distance))
identical(taildist1$n,taildist2$not_cancelled)
```

- - - 

*4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?*

__ANSWER:__ 

```{r 5_6_7_EX4, warning=FALSE, message=FALSE}
# To answer this question, we first need to create a data set of relevant delay statistics:
delays <- flights %>%
        group_by(year, month, day) %>%
        summarize(count = n(),
                  cancelled = sum(is.na(dep_delay)),
                  avg_dep_delay = mean(mean(dep_delay,na.rm=TRUE)),
                  avg_arr_delay = mean(mean(arr_delay,na.rm=TRUE)))
# Now we can use ggplot to look into delay patterns
plot1 <- ggplot(data = delays, mapping = aes(x = cancelled, y = avg_dep_delay)) +
  geom_point(alpha = 1/3) +
  geom_smooth(se = FALSE)
plot2 <- ggplot(data = delays, mapping = aes(x = cancelled, y = avg_arr_delay)) +
  geom_point(alpha = 1/3) +
  geom_smooth(se = FALSE)
grid.arrange(plot1, plot2, ncol=2)
# We can see that for the most part, days with high average departure or arrival delays have higher numbers of cancelled flights. However, the relationship isn't linear and there are some clear outliers.
```

- - - 

*6. What does the sort argument to count() do. When might you use it?*

__ANSWER:__ The sort argument allows you to sort output in descendiing order. This could be useful when looking for extremes (either highest or lowest frequencies).

####10.5 Exercises:####

*1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).*

__ANSWER:__ R will identify tibbles when you use the head() function, or if you try to print the entire dataset. 

- - - 

*2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?*

__ANSWER:__ 

```{r 10_5_EX2, warning=FALSE, message=FALSE}
# Two equivalent data sets, one as a data.frame and one as a tibble
df <- data.frame(abc = 1, xyz = "a")
df_tbl <- tibble(
  abc = 1, 
  xyz = "a"
  )

# If you mistype a column name in a data.frame, R will still return a value if the name is close, whereas a tibble will not:
df$x
df_tbl$x

# If you specify the data.frame column(s) correctly, R will return the values but a tibble will give you more useful information like the column type (character) and dimensions:
df[, "xyz"]
df_tbl[, "xyz"]

df[, c("abc", "xyz")]
df_tbl[, c("abc", "xyz")]
```

- - - 

*3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?*

__ANSWER:__ You can either use [1] or [[1]], where 1 corresponds to the column containing var.

```{r 10_5_EX3, warning=FALSE, message=FALSE}
df <- tibble(var <- "mpg")
df[[1]]
df[1]
```

- - - 

*6. What option controls how many additional column names are printed at the footer of a tibble?*

__ANSWER:__ The n_extra argument can be added to a print() statement to control the number of extra columns listed at the bottom of the footer.

- - - 

####12.3.3 Exercises:####

*2. Why does this code fail?*

```{r 12_3_3_EX2.1, warning=FALSE, message=FALSE, eval=FALSE}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n
```

__ANSWER:__ The code fails because the two year columns are numbers. In order to make the code work you have to surround them in quotes like this:

```{r 12_3_3_EX2.2, warning=FALSE, message=FALSE}
table4a %>% 
  gather('1999', '2000', key = "year", value = "cases")
```

- - - 

*3. Why does spreading this tibble fail? How could you add a new column to fix the problem?*

__ANSWER:__ Spreading this tibble fails because there are two different ages for the same person. You could add a new column (let's say year) to differentiate between the two, like this:

```{r 12_3_3_EX3.1, warning=FALSE, message=FALSE}
people <- tribble(
  ~name,             ~key,    ~value, ~year,
  #-----------------|--------|------|------
  "Phillip Woods",   "age",       45, 2000,
  "Phillip Woods",   "height",   186, 2000,
  "Phillip Woods",   "age",       50, 2005,
  "Jessica Cordero", "age",       37, 2000,
  "Jessica Cordero", "height",   156, 2000
)
```

```{r 12_3_3_EX3.2, warning=FALSE, message=FALSE}
people %>%
    spread(key, value)
```
- - - 

*4. Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?*

__ANSWER:__ 

```{r 12_3_3_EX4.1, warning=FALSE, message=FALSE}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```

```{r 12_3_3_EX4.2, warning=FALSE, message=FALSE}
# You can tidy the tibble by using gather()
preg %>% 
  gather(male, female, key = "gender", value = "count")
```
- - - 

####12.4.3 Exercises:####

*1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.*

__ANSWER:__ extra can be used when there are too many pieces, like in the first example. fill can be used when there are not enough pieces, like in the second example.

```{r 12_4_3_EX1, warning=FALSE, message=FALSE}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left")
```

- - - 

*2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?*

__ANSWER:__ The remove argument removes the input column (in the example below, this is the rate column). If set to false, it does not remove it. This could be useful if you want to confirm that the separation worked as you expected.

```{r 12_4_3_EX2, warning=FALSE, message=FALSE}
# remove set to TRUE (default value)
table3 %>% 
  separate(rate, into = c("cases", "population"), convert = TRUE, remove = TRUE)
# remove set to FALSE 
table3 %>% 
  separate(rate, into = c("cases", "population"), convert = TRUE, remove = FALSE)
```

- - - 

####Data Import/Export Exercise:####

Follow these steps:

* Download the baby_names.txt file from Canvas which is in the Homework 4 assignment section.
* Load this file into R correctly and take a glimpse of the output.
* Export this file as a csv file and call it ‘baby_names.csv’.
* Reload the baby_names.csv file and take another glimpse.
* Show all of your code and the output. There should be two data import lines of code, one data export line of code, and two glimpses of the data.

__ANSWER:__ 
```{r import_export, warning=FALSE, message=FALSE}
#Load baby_names.txt file into R and take a glimpse of the output
baby_names_path <- 'C:/Users/Eeslami/Desktop/UCB Courses/X415.2 - Introduction to Data Science/compscix-415-2-assignments/baby_names.txt'
baby_names_df <- read_delim(file = baby_names_path, delim = '|')
glimpse(baby_names_df)

#Export baby_names.txt file as a csv file and call it ‘baby_names.csv’
write_delim(baby_names_df, delim = ',', path = 'baby_names.csv')

#Reload the baby_names.csv file and take another glimpse.
baby_names_df2 <- read.csv(file = 'baby_names.csv')
glimpse(baby_names_df2)
```

- - - 

