---
author: Esa Eslami
date: 7/22/2018
title: "COMPSCIX 415.2 Homework 7"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
---

```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(broom)

```

##__Homework 7 Assignments:__##

### Exercise 1

*Load the train.csv dataset into R. How many observations and columns are there?*

__ANSWER:__ There are 1,460 observations (rows) and 81 variables (columns).

```{r ex1_q1, warning=FALSE, message=FALSE, eval=TRUE}
#Set path
train_df_path <- 'C:/Users/Eeslami/Desktop/UCB Courses/X415.2 - Introduction to Data Science/compscix-415-2-assignments/train.csv'
#Load file 
train_df <- read_csv(file = train_df_path)
#Get dimensions of data frame
train_df %>%
  as_tibble() %>%
  dim_desc()
```

* * *
### Exercise 2

*Normally at this point you would spend a few days on EDA, but for this homework we will do some very basic EDA and get right to fitting some linear regression models.*

*Our target will be SalePrice.*

* Visualize the distribution of SalePrice. 

__ANSWER:__ 

```{r ex2_q1, warning=FALSE, message=FALSE, eval=TRUE}
library(scales)
train_df %>% ggplot(aes(x = SalePrice)) +
  geom_histogram()+ 
  labs(x="Sale Price") +
  scale_x_continuous(label=dollar_format()) +
  theme_bw()
```

* Visualize the covariation between SalePrice and Neighborhood. 

__ANSWER:__ 

```{r ex2_q2, warning=FALSE, message=FALSE, eval=TRUE}
train_df %>% ggplot(aes(x = SalePrice, y = reorder(Neighborhood, SalePrice, FUN = median))) +
  geom_point(alpha = .8) +
  labs(x="Sale Price", y = "Neighborhood") +
  scale_x_continuous(label=dollar_format()) +
  theme_bw()
```

* Visualize the covariation between SalePrice and OverallQual. 

__ANSWER:__ 

```{r ex2_q3, warning=FALSE, message=FALSE, eval=TRUE}
train_df %>% ggplot(aes(x = SalePrice, y = OverallQual)) +
  geom_point(alpha = .8) +
  labs(x="Sale Price") +
  scale_x_continuous(label=dollar_format()) +
  theme_bw()
```

* * *
### Exercise 3

*Our target is called SalePrice. First, we can fit a simple regression model consisting of only the intercept (the average of SalePrice). Fit the model and then use the broom package to:*

* take a look at the coefficient, 

__ANSWER:__ 

```{r ex3_q1, warning=FALSE, message=FALSE, eval=TRUE}

```

* compare the coefficient to the average value of SalePrice, and 

__ANSWER:__ 

```{r ex3_q2, warning=FALSE, message=FALSE, eval=TRUE}

```

* take a look at the R-squared. 

__ANSWER:__ 

```{r ex3_q3, warning=FALSE, message=FALSE, eval=TRUE}

```

* * *

### Exercise 4

*Now fit a linear regression model using GrLivArea, OverallQual, and Neighborhood as the features. Don’t forget to look at data_description.txt to understand what these variables mean. Ask yourself these questions before fitting the model:*

* What kind of relationship will these features have with our target? 

__ANSWER:__ 

```{r ex4_q1, warning=FALSE, message=FALSE, eval=TRUE}

```

* Can the relationship be estimated linearly? 

__ANSWER:__ 

```{r ex4_q2, warning=FALSE, message=FALSE, eval=TRUE}

```

* Are these good features, given the problem we are trying to solve? 

__ANSWER:__ 

```{r ex4_q3, warning=FALSE, message=FALSE, eval=TRUE}

```

*After fitting the model, output the coefficients and the R-squared using the broom package. Answer these questions:*

* How would you interpret the coefficients on GrLivArea and OverallQual? 

__ANSWER:__ 

```{r ex4_q4, warning=FALSE, message=FALSE, eval=TRUE}

```

* How would you interpret the coefficient on NeighborhoodBrkSide? 

__ANSWER:__ 

```{r ex4_q5, warning=FALSE, message=FALSE, eval=TRUE}

```

* Are the features significant? 

__ANSWER:__ 

```{r ex4_q6, warning=FALSE, message=FALSE, eval=TRUE}

```

* Are the features practically significant? 

__ANSWER:__ 

```{r ex4_q7, warning=FALSE, message=FALSE, eval=TRUE}

```

* Is the model a good fit? 

__ANSWER:__ 

```{r ex4_q8, warning=FALSE, message=FALSE, eval=TRUE}

```
* * *

### Exercise 6

*One downside of the linear model is that it is sensitive to unusual values because the distance incorporates a squared term. Fit a linear model to the simulated data below (use y as the target and x as the feature), and look at the resulting coefficients and R-squared. Rerun it about 5-6 times to generate different simulated datasets. What do you notice about the model’s coefficient on x and the R-squared values?*

```{r ex6_q1, warning=FALSE, message=FALSE, eval=FALSE}
sim1a <- tibble(
x = rep(1:10, each = 3),
y = x * 1.5 + 6 + rt(length(x), df = 2)
)
```

__ANSWER:__ 

