---
author: Esa Eslami
date: 7/01/2018
title: "COMPSCIX 415.2 Homework 4"
output: html_document
self-contained: true
---

```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(nycflights13)
```

* * * 

####__Homework 4 Assignments:__####
__Complete these Exercises in R for Data Science:__  

Section 5.6.7: #2, #4 and #6 only. Extra Credit: Do #5
Section 10.5: #1, #2, #3 and #6 only
Section 12.3.3: #2, #3 and #4 only
Section 12.4.3: #1 and #2 only

* * * 

####5.6.7 Exercises:####

*2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).*

__ANSWER:__ 

```{r 5_6_7_EX2, warning=FALSE, message=FALSE}
# First we need to create non_cancelled
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
# Now run not_cancelled %>% count(dest) to see what it looks like
not_cancelled %>% count(dest)
# We can get this same output using summarize() like this:
flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
  group_by(dest) %>%
  summarize(not_cancelled = sum(!is.na(dep_delay)))
# Confirm that the output is identical:
countdest1 <- not_cancelled %>% count(dest)
countdest2 <- flights %>%
        filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
        group_by(dest) %>%
        summarize(not_cancelled = sum(!is.na(dep_delay)))
identical(countdest1$n,countdest2$not_cancelled)
# Similarly, we can use summarize() to get the same results from not_cancelled %>% count(tailnum, wt = distance) like this:
taildist1 <- not_cancelled %>% count(tailnum, wt = distance)
taildist2 <- flights %>%
        filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
        group_by(tailnum) %>%
        summarize(not_cancelled = sum(distance))
identical(taildist1$n,taildist2$not_cancelled)
```

- - - 

*4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?*

__ANSWER:__ 

```{r 5_6_7_EX4, warning=FALSE, message=FALSE}

```

- - - 

*6. What does the sort argument to count() do. When might you use it?*

__ANSWER:__ 

```{r 5_6_7_EX6, warning=FALSE, message=FALSE}

```

####10.5 Exercises:####

*1. ?*

__ANSWER:__ 

```{r 10_5_EX1, warning=FALSE, message=FALSE}

```

- - - 

*2. ?*

__ANSWER:__ 

```{r 10_5_EX2, warning=FALSE, message=FALSE}

```

- - - 

*3. ?*

__ANSWER:__ 

```{r 10_5_EX3, warning=FALSE, message=FALSE}

```

- - - 

*6. ?*

__ANSWER:__ 

```{r 10_5_EX6, warning=FALSE, message=FALSE}

```

- - - 
####12.3.3 Exercises:####

*2. ?*

__ANSWER:__ 

```{r 12_3_3_EX2, warning=FALSE, message=FALSE}

```

- - - 

*3. ?*

__ANSWER:__ 

```{r 12_3_3_EX3, warning=FALSE, message=FALSE}

```

- - - 

*4. ?*

__ANSWER:__ 

```{r 12_3_3_EX4, warning=FALSE, message=FALSE}

```

- - - 

####12.4.3 Exercises:####

*1. ?*

__ANSWER:__ 

```{r 12_4_3_EX1, warning=FALSE, message=FALSE}

```

- - - 

*2. ?*

__ANSWER:__ 

```{r 12_4_3_EX2, warning=FALSE, message=FALSE}

```

- - - 

