---
author: Esa Eslami
date: 6/18/2018
title: "COMPSCIX 415.2 Homework 2"
output: html_document
self-contained: true
---

```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
```

* * * 

####__Homework 2 Assignments:__####
__Complete these Exercises in R for Data Science:__  
1. Section 3.2.4: all exercises  
2. Section 3.3.1: all exercises  
3. Section 3.5.1: #4 and #5 only  
4. Section 3.6.1: #1-5. Extra Credit: Do #6  
5. Section 3.7.1: #2 only  

* * * 

####3.2.4 Exercises:####

*1. Run ggplot(data = mpg). What do you see?*
__ANSWER:__ ggplot(data = mpg) produces an empty graph:
```{r 3_2_4_exercises1, warning=FALSE, message=FALSE}
ggplot(data = mpg)
```
- - - 

*2. How many rows are in mpg? How many columns?*

```{r 3_2_4_exercises2, warning=FALSE, message=FALSE, echo=FALSE, eval=FALSE}
glimpse(mpg)
```

__ANSWER:__ There are 234 observations (rows) and 11 variables (columns) in the mpg data set.

- - - 

*3. What does the drv variable describe? Read the help for ?mpg to find out.*

__ANSWER:__ The drv variable describes whether a car is front-wheel drive (coded as f), rear wheel drive (coded as r), or 4wd (coded as 4).

- - - 

*4. Make a scatterplot of hwy vs cyl.*

__ANSWER:__ 
```{r 3_2_4_exercises4, warning=FALSE, message=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

- - - 

*5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?*

__ANSWER:__ class and drv are both categorical variables, so a scatterplot does not give a useful visual in the same way that it would were the variables numeric.
```{r 3_2_4_exercises5, warning=FALSE, message=FALSE, echo=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

- - - 

