---
author: Esa Eslami
date: 6/11/2018
title: "COMPSCIX 415.2 Homework 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL: [https://github.com/esaeslami/compscix-415-2-assignments](https://github.com/esaeslami/compscix-415-2-assignments)

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)

data("WorldCities")
glimpse(WorldCities)
```

*__QUESTION 1:__ How many observations and variables are there in this data set? What are some of the variable names? Type up your answer in your document using complete sentences.*

**ANSWER 1:** There are 23,018 observations and 10 variables in the "WorldCities" data set. The 10 variable names are: code, name, latitude, longitude, country, countryRegion, population, regionCode, region, and date.

__Extract parts of the data (*EXAMPLE*)__

```{r Q1 example, warning=FALSE, message=FALSE}
WorldCities <- head(WorldCities, 200) # 200 rows
country_col <- WorldCities$country
unique(country_col)
```

*__QUESTION 2:__ There is a column called region in the data set. Can you extract this and show only the unique values?*

**ANSWER 2:** 
```{r Q2, warning=FALSE, message=FALSE}
region_col <- WorldCities$region
unique(region_col)
```