---
author: Esa Eslami
date: 6/18/2018
title: "COMPSCIX 415.2 Homework 2"
output: html_document
self-contained: true
---

```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
```

* * * 

####__Homework 2 Assignments:__####
__Complete these Exercises in R for Data Science:__  
1. Section 3.2.4: all exercises  
2. Section 3.3.1: all exercises  
3. Section 3.5.1: #4 and #5 only  
4. Section 3.6.1: #1-5. Extra Credit: Do #6  
5. Section 3.7.1: #2 only  

* * * 

####3.2.4 Exercises:####

*1. Run ggplot(data = mpg). What do you see?*

__ANSWER:__ ggplot(data = mpg) produces an empty graph:
```{r 3_2_4_exercises1, warning=FALSE, message=FALSE}
ggplot(data = mpg)
```

- - - 

*2. How many rows are in mpg? How many columns?*

```{r 3_2_4_exercises2, warning=FALSE, message=FALSE, echo=FALSE, eval=FALSE}
glimpse(mpg)
```

__ANSWER:__ There are 234 observations (rows) and 11 variables (columns) in the mpg data set.

- - - 

*3. What does the drv variable describe? Read the help for ?mpg to find out.*

__ANSWER:__ The drv variable describes whether a car is front-wheel drive (coded as f), rear wheel drive (coded as r), or 4wd (coded as 4).

- - - 

*4. Make a scatterplot of hwy vs cyl.*

__ANSWER:__ 
```{r 3_2_4_exercises4, warning=FALSE, message=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

- - - 

*5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?*

__ANSWER:__ class and drv are both categorical variables, so a scatterplot does not give a useful visual in the same way that it would were the variables numeric.

```{r 3_2_4_exercises5, warning=FALSE, message=FALSE, echo=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

- - - 

####3.3.1 Exercises:####

*1. What’s gone wrong with this code? Why are the points not blue?*
```{r 3_3_1_exercises1_1, warning=FALSE, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

__ANSWER:__ color needs to be outside of the aes(), like this: 
```{r 3_3_1_exercises1_2, warning=FALSE, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

- - - 

*2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?*

__ANSWER:__ manufacturer, model, trans, drv, fl, and class are categorical. displ, year, cyl, cty, and hwy are continuous. You can see this information with glimpse():
```{r 3_3_1_exercises2, warning=FALSE, message=FALSE}
glimpse(mpg)
```

- - - 

*3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?*

__ANSWER:__ ggplot() will give an error if you try to map a continuous variable to shape. However, it will allow you to map continuous variables to color and size:
```{r 3_3_1_exercises3, warning=FALSE, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cyl))
```

Color and size behave differently for categorical vs. continuous variables in that they are easier to distinguish for categorical variables. For continuous variables, color and size vary more at extremes.

- - - 

*4. What happens if you map the same variable to multiple aesthetics?*

__ANSWER:__ ggplot() will not stop you from mapping the same variable to multiple aesthetics:
```{r 3_3_1_exercises4, warning=FALSE, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = cty, color = cty, size = cty))
```

- - - 

*5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)*

__ANSWER:__ stroke increases the thickness of the border around dots in geom_point. It works with all shapes, but can make it difficult to distinguish between some shapes:
```{r 3_3_1_exercises5, warning=FALSE, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = cty, shape = class, stroke = 3))
```

- - - 

*6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?*

__ANSWER:__ ggplot() allows you to map conditions to aesthetics, like this:
```{r 3_3_1_exercises6, warning=FALSE, message=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = cty, color = displ < 5))
```


