---
author: Esa Eslami
date: 6/22/2018
title: "COMPSCIX 415.2 Homework 3"
output: html_document
self-contained: true
---

```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(nycflights13)
```

* * * 

####__Homework 3 Assignments:__####
__Complete these Exercises in R for Data Science:__  

1. Section 3.8.1: all exercises
2. Section 3.9.1: #2 and #4 only
3. Section 4.4: #1 and #2 only
4. Section 5.2.4: #1, #3 and #4 only (use nycflights13 package)
5. Section 5.4.1: #1 and #3 only 

* * * 

####3.8.1 Exercises:####

*1. What is the problem with this plot? How could you improve it?*

__ANSWER:__ Like the hwy by displ example plot in this section, plotting hwy against cty results in a lot of overlapping points. You could improve the plot by using the position = "jitter" option, like this:

```{r 3_8_1_exercises1, warning=FALSE, message=FALSE}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

- - - 

*2. What parameters to geom_jitter() control the amount of jittering?*

__ANSWER:__ The width and height parameters for geom_jitter() control the amount of jittering.

- - - 

*3. Compare and contrast geom_jitter() with geom_count().*

__ANSWER:__ geom_jitter() is a useful shortcut when you know you will have overlapping points and don't want to add the position = "jitter" option to geom_point(). geom_count, on the other hand, visualizes the amount of overlap at each position by increasing the area of points with overlap. geom_count() therefore is more useful when you care about highlighting overlap, while geom_jitter() is more of a programming shortcut.

- - - 

*4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.*

__ANSWER:__ The default position adjustment for geom_boxplot() is position = "dodge":

```{r 3_8_1_exercises4, warning=FALSE, message=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = cty)) + 
  geom_boxplot(aes(color = drv), position = "dodge")
```

- - - 

####3.9.1 Exercises:####

*2. What does labs() do? Read the documentation.*

__ANSWER:__ labs() allows you to define the x label, y label, and title(s) of a plot.

- - - 

*4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?*

__ANSWER:__ The plot below shows that as city miles per gallon increase so does highway miles per gallon. The coord_fixed() option is used to make the graphical display of the x axis proportional to the y axis. The geom_abline() option adds a fitted line to the graph to demonstate the linear relationship of the two variables.

```{r 3_9_1_exercises4, warning=FALSE, message=FALSE}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point()  +
  geom_abline() +
  coord_fixed()
```

- - - 

####4.4 Exercises:####

*1. Why does this code not work?*

__ANSWER:__ The code doesn't work because the i in my_variable is replaced by a 1 (or something that isn't the letter i) in my_varıable.

- - - 

*2. Tweak each of the following R commands so that they run correctly:*

__ANSWER:__ 

```{r 4_4_exercises2, warning=FALSE, message=FALSE}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)
filter(diamonds, carat > 3)
```

- - - 

####5.2.4 Exercises:####

*1. Find all flights that...*

__ANSWER:__ 

```{r 5_2_4_exercises1, warning=FALSE, message=FALSE}
glimpse(flights)
#1. Had an arrival delay of two or more hours
filter(flights, dep_delay >= 2)
#2. Flew to Houston (IAH or HOU)
filter(flights, dest == "HOU")
#3. Were operated by United, American, or Delta
filter(flights, carrier %in% c("UA", "AA", "DL"))
#4. Departed in summer (July, August, and September)
filter(flights, month %in% c(7:9))
#5. Arrived more than two hours late, but didn’t leave late
filter(flights, arr_delay > 2, dep_delay == 0)
#6. Were delayed by at least an hour, but made up over 30 minutes in flight
filter(flights, arr_delay <= -0.5, dep_delay >= 1)
#7. Departed between midnight and 6am (inclusive)
filter(flights, dep_time %in% c(2400,0:600))

```

- - - 

*3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?*

__ANSWER:__ There are 8,255 flights with missing depature times. Variables that are dependant on a departure (departure delay, arrival time, arrival delay, and air time) are also all missing for these flights.

```{r 5_2_4_exercises3, warning=FALSE, message=FALSE}
#how many flights with a missing dep_time?
count(filter(flights, is.na(dep_time)))
#What other variables are missing for these flights?
filter(flights, is.na(dep_time))
```

- - - 

*4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)*

__ANSWER:__ NA ^ 0 is not missing because any value to the zero is 1. NA | TRUE is not missing TRUE is an option. NA & FALSE is not missing because NA is not FALSE. I think the general rule is that NA is treated like an unknown value, rather than a NULL or empty value.

- - - 

####5.4.1 Exercises:####

*1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights*

__ANSWER:__ 

```{r 5_4_1_exercises1, warning=FALSE, message=FALSE}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, dep_time:arr_delay, -sched_dep_time, -sched_arr_time)
select(flights, starts_with("dep"), starts_with("arr"))
```

- - - 

*3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?*

__ANSWER:__ one_of() is useful here because if you try to use the vars vector to select the column names alone, you will get an error. If you use one_of() to read in the names this will work in R. 

```{r 5_4_1_exercises3, warning=FALSE, message=FALSE}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))

```

- - - 













