---
author: Esa Eslami
date: 6/22/2018
title: "COMPSCIX 415.2 Homework 3"
output: html_document
self-contained: true
---

```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(nycflights13)
```

* * * 

####__Homework 3 Assignments:__####
__Complete these Exercises in R for Data Science:__  

1. Section 3.8.1: all exercises
2. Section 3.9.1: #2 and #4 only
3. Section 4.4: #1 and #2 only
4. Section 5.2.4: #1, #3 and #4 only (use nycflights13 package)
5. Section 5.4.1: #1 and #3 only 

* * * 

####3.8.1 Exercises:####

*Question 1.*

__ANSWER:__ Like the hwy by displ example plot in this section, plotting hwy against cty results in a lot of overlapping points. You could improve the plot by using the position = "jitter" option, like this:

```{r 3_8_1_exercises1, warning=FALSE, message=FALSE}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

- - - 

*Question 2. *

__ANSWER:__ The width and height parameters for geom_jitter() control the amount of jittering.

- - - 

*Question 3. *

__ANSWER:__ geom_jitter() is a useful shortcut when you know you will have overlapping points and don't want to add the position = "jitter" option to geom_point(). geom_count, on the other hand, visualizes the amount of overlap at each position by increasing the area of points with overlap. geom_count() therefore is more useful when you care about highlighting overlap, while geom_jitter() is more of a programming shortcut.

- - - 

*Question 4. *

__ANSWER:__ The default position adjustment for geom_boxplot() is position = "dodge":

```{r 3_8_1_exercises4, warning=FALSE, message=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = cty)) + 
  geom_boxplot(aes(color = drv), position = "dodge")
```

- - - 

####3.9.1 Exercises:####

*2. *

__ANSWER:__ labs() allows you to define the x label, y label, and title(s) of a plot.

- - - 

*4. *

__ANSWER:__ The plot below shows that as city miles per gallon increase so does highway miles per gallon. The coord_fixed() option is used to make the graphical display of the x axis proportional to the y axis. The geom_abline() option adds a fitted line to the graph to demonstate the linear relationship of the two variables.

```{r 3_9_1_exercises4, warning=FALSE, message=FALSE}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point()  +
  geom_abline() +
  coord_fixed()
```

- - - 

####4.4 Exercises:####

*1. *

__ANSWER:__ The code doesn't work because the i in my_variable is replaced by a 1 (or something that isn't the letter i) in my_varıable.

- - - 

*2. *

__ANSWER:__ 

```{r 4_4_exercises2, warning=FALSE, message=FALSE}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)
filter(diamonds, carat > 3)
```

- - - 

####5.2.4 Exercises:####

*1. *

__ANSWER:__ 

```{r 5_2_4_exercises1, warning=FALSE, message=FALSE}

```

- - - 

*3. *

__ANSWER:__ 

```{r 5_2_4_exercises3, warning=FALSE, message=FALSE}

```

- - - 

*4. *

__ANSWER:__ 

```{r 5_2_4_exercises4, warning=FALSE, message=FALSE}

```

- - - 

####5.4.1 Exercises:####

*1. *

__ANSWER:__ 

```{r 5_4_1_exercises1, warning=FALSE, message=FALSE}

```

- - - 

*3. *

__ANSWER:__ 

```{r 5_4_1_exercises3, warning=FALSE, message=FALSE}

```

- - - 













